---
title: "Simple Regress"
author: "Claudia Cortez"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
For model length ~ neck,agrested r^2=0.75

```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)    
```
  adjuste r^2 for the age model is 0.51, this is less than the adjuste r^2 for neck, so this model is not as good as the neck model 

## length based on chest

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)    
```
  The model length based on chest is better than adjusted $r^2$ for the age model (0.51)and adjested r^2 for neck model (0.75),because r^2 for chest is (0.79). 


### length based on weight 
```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)    
```
  The model length based weight is not as good as the model - length based on chest because the adjusted r-squared for length based on weight is lower (0.74) than the adjuested r-squared for chest (0.79).

### length based on sex 
```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)    
```
  This model is not better than the chest model beause the rsqure for length based on sex is (-0.0) this means its lower than chest r-squared (0.79).

### length based on headlen
```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)    
```
This model length based on headlen based on leadlen is way better than length based on chest becasue the r-squared for chest is (0.78)and whish is lower than the r-squard for headlen which was (0.84).  



### length based on headwth
```{r}
headwthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwthModel)
summary.lm(headWthModel)    
```
This model is not better than the model headlen because the r^2 for headwith is lower (0.53) and the headlen was (0.84)which is higher 

### Conclusion 
The greates model value was lenght~ Headlen the rquare was 0.84. 
 

## Compare all models  that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), by RSE (residual standard error) where smaller is better. 

